<FONT SIZE=4, COLOR="#8272E9"><CENTER><B>
<P>
Mohan , Sagar, Prateek
<BR>
ALY6015 - Intermediate Analytics
<BR>
Northeastern University
<BR>
Dr. Ji-Young Yun
<BR>
Date of submission: February 2021
<BR>
Module 1
</CENTER>
</FONT>
</B>

<h3> <b>Introduction: <b></h3>
```{r}
install.packages("ggcorrplot")

install.packages("ggplot2")
```


```{r}
library("ggpubr")
library("readxl")
library("dplyr")
library("readr")
library("knitr")
library("kableExtra")
library("tidyverse")
library("vcd")
library("lessR")
library("RColorBrewer")
library(lattice)
library("corrplot")
library(Hmisc)
library("ggplot2")
library(ggcorrplot)
library("car")
library("caret")
library("ggplot2")
library(gridExtra)
library("zoo")
library("ISLR")
library(pROC)

library(tidyverse)
library(nnet)
library(car)

```


```{r}
rawdata = read_csv("../Intermediate Analytics/DeAP (1).csv")

rawdata

```




```{r}
glimpse(rawdata)
```


```{r}
colnames(rawdata)
```

# Removing the unnecessary columns from the data.

```{r}

drop_cols = c("StartDate" ,"EndDate" ,"Status" , "IPAddress" ,"Progress", "Duration (in seconds)","Finished","RecordedDate","ResponseId", "RecipientLastName",  "RecipientFirstName","RecipientEmail","ExternalReference" ,"LocationLatitude" ,"LocationLongitude","DistributionChannel","UserLanguage")

sorted_data = rawdata[,!names(rawdata)%in%drop_cols]                
sorted_data

```

#Renaming the column names into an meaningful variables

```{r}
colnames(sorted_data) = c("gender","age","profession","other","avgworkhour","sleep","activity","labour/sittingjob","spare_time","screentime","height","weight","hobbies",
                          "steps","waterintake","depression","sleep_disorder","smoking","drinking")
```

```{r}
head(sorted_data)
tail(sorted_data)
```
#Finding and  Handling Nans 


```{r}
null_count = data.frame(map(sorted_data, ~sum(is.na(.))))


a =apply(null_count[,-1],1,function(x) names(null_count[,-1])[which(x>0)])

t(null_count[a[,]])
```

From the above output we can see that other column has majority of nulls so we drop the "other" column.

```{r}
drop_cols = c("other")
sorted_data = sorted_data[,!names(sorted_data)%in%drop_cols]                
head(sorted_data)
```
# removing the first two rows in all columns as the data is messy


```{r}
sorted_data = sorted_data %>% slice(-c(1,2))
```

```{r}
sorted_data
```


```{r}
clean_data = na.omit(sorted_data)
clean_data
```

#duplicate values in the dataset

```{r}

cat("Total number of duplicates in the dataset are",sum(duplicated(clean_data)))
```

# cleaning unformatted data 
```{r}
library(tidyr)

#df <- data.frame(clean_data = c("height"))

clean_data$height <- as.integer(gsub('[a-zA-Z]', '', clean_data$height))

head(clean_data)
```

```{r}
head(clean_data)
```

# changing the datatypes of the numerical values
```{r}
numerical_cols = c( "age","avgworkhour","sleep","screentime","height","weight", "steps","waterintake") 

clean_data[, c( "age","avgworkhour","sleep","screentime","height","weight", "steps","waterintake")] <- sapply(clean_data[, c( "age","avgworkhour","sleep","screentime","height","weight", "steps","waterintake")], as.numeric)

```





#We are picking five random numerical values and analysing them.
```{r}
numerical_analysis = matrix(c(summary(clean_data$steps),summary(clean_data$screentime),summary(clean_data$age),summary(clean_data$avgworkhour),summary(clean_data$sleep),summary(clean_data$weight),summary(clean_data$height),summary(clean_data$waterintake)),nrow = 8, byrow = TRUE)
rownames(numerical_analysis) = c("Steps","Screen Time","Age","Avg Work Hour","Sleep", "Weight","Height","Water Intake")
colnames(numerical_analysis) = c("Minimum","1st","Median","Mean","3rd Quartile","Maximum")

kable(x = numerical_analysis, format = 'html',
      caption = "Summary of  DeAP data") %>%
   kable_material_dark(html_font = "Cambria") %>%
  kable_styling(font_size = 15) %>%
gsub("font-size: initial !important;",
"font-size: 10 pt !important;",
.)%>%
row_spec(0, bold = TRUE, color = "#33FFE3", background = "#581845", font_size = 14) %>%
column_spec(1, bold = TRUE, border_right = TRUE, color = "#810033", background = "#DAF7A6") 


```

#outlier detection:

```{r}
a = boxplot(clean_data$sleep, main="sleep",
    ylab="Sleep time in hours", col="green",
border="blue")

boxplot
```

```{r}
a = boxplot(clean_data$screentime, main="sleep",
    ylab="Sleep time in hours", col="green",
border="blue")

a$out

```

```{r}



counts <- table(clean_data$depression)

x= barplot(counts,
        main="Analysis of dependant variable in the dataset",
        xlab="Depression ? ", ylab="Frequency",col=brewer.pal(5,"Set1"),border="white",
ylim=c(0,20))
text(x,counts+2,labels = as.character(counts))

        
```
```{r}
#x <- table(na.omit(data_cat$Alley))
x <- table(clean_data$depression)
PieChart(x, hole = 0, values = "%", main = "Analysis of  dependant variable (depression) in the dataset ")
```




#
```{r}
dat <- data.frame(table(clean_data$smoking,clean_data$depression))
names(dat) <- c("smoking","depression","Count")

ggplot(data=dat, aes(x=depression, y=Count, fill=smoking)) + geom_bar(stat="identity")




```

```{r}
dat <- data.frame(table(clean_data$drinking,clean_data$depression))
names(dat) <- c("drinking","depression","Count")

ggplot(data=dat, aes(x=depression, y=Count, fill=drinking)) + geom_bar(stat="identity")



```
```{r}
p <- ggplot(clean_data, aes(x=depression, y=screentime, fill=depression)) + # fill=name allow to automatically dedicate a color for each group
  geom_violin()
p
```


# total steps vs depression
```{r}
boxplot(steps~depression,
data=clean_data,
main="Different boxplots for each type of Depression",
xlab="depression",
ylab=" total steps per day",
col="maroon",
border="blue"
)
```


```{r}
clean_data$depression <- as.factor(clean_data$depression)
clean_data$sleep_disorder <- as.factor(clean_data$sleep_disorder)
#clean_data$profession <- as.factor(clean_data$profession)

clean_data$depression <- relevel(clean_data$depression, "No") # Set Reference level to no response...
clean_data$sleep_disorder<- relevel(clean_data$sleep_disorder, "No")
#clean_data$profession = relevel(clean_data$profession, "Student")
```

# Split data into train and test sets

```{r}
set.seed(48)
trainIndex = createDataPartition(clean_data$depression, p =0.70,list = FALSE)

clean_data_train = clean_data[trainIndex,]
clean_data_test  = clean_data[-trainIndex,]

head(clean_data_train)
```
```{r}

```

# Running model with all predictors 
```{r}

Model1 <- multinom(depression ~ ., data = clean_data_train)
```

```{r}
 a = summary(Model1)
a
```


```{r}
colnames(clean_data)
```



```{r warning=FALSE}
model2 <- multinom(depression ~ gender + age + profession + avgworkhour + sleep + activity + spare_time + screentime + height + weight  + steps + waterintake  + sleep_disorder  + smoking + drinking   , data = clean_data_train)
```



```{r}
 b = summary(model2)
b
```

```{r}
b $ call
```

```{r}
model3 <- multinom(depression ~ age + avgworkhour + sleep + activity + screentime + weight  + steps + waterintake + sleep_disorder  + smoking + drinking   , data = clean_data_train)
```


```{r}
c = summary(model3)
c
```



#Confusion matrix






```{r}
#test set predictions

predicted_value = predict(model3, newdata = clean_data_test)

#predicted_classes_min = as.factor(ifelse(probabilites_test >= 0.05,"Yes", "No"))

#confusionMatrix(predicted_classes_min,clean_data_test$depression, positive = 'Yes')

expected_value = clean_data_test$depression

model_conf =  confusionMatrix(data=predicted_value, reference = expected_value)

```

```{r}
model_conf
```
```{r}
clean_data_test$depression
```
```{r}
predicted_value <- as.character(predicted_value)
predicted_value[predicted_value == "Yes"] = 2      # Replace "Yes" by 2
predicted_value[predicted_value == "Maybe"] = 1
predicted_value[predicted_value == "No"] = 0
#predicted_value_num <- as.numeric(predicted_value)

predicted_value_num = c(2,0,1,0)
expected_value_num = c(1,0,0,1)

```


```{r}
model3_roc  = roc(expected_value_num, predicted_value_num)

plot(model1_roc,col = "blue",ylab = "Sensitivity - TP Rate", xlab = "Specificity - FP Rate")
```

```{r}
model3_auc  = auc(model3_roc)
model3_auc
```




















